---
title: "bootstrap-confidence-intervals"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bootstrap-confidence-intervals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# command line instruction to build the vignette: 
# devtools::build_rmd("vignettes/bootstrap-confidence-intervals.Rmd")

```

# Bootstrap Confidence Intervals

```{r setup}
library(tidyverse)
library(BehavioralDataAnalysis)
library(MASS)
```

## Introduction and basic use cases

### Introduction

When running data analyses, we're often interested in determining how accurate they would be on new data, for example extrapolating from the past to the future. The traditional approach is to assume that the existing data and any new data both follow the same statistical distribution, whose parameters we then estimate. The Bootstrap is an alternative approach, which relies on computer simulations instead. You can learn more about the Bootstrap in my book [Behavioral Data Analysis with R and Python](https://www.amazon.com/Behavioral-Data-Analysis-Python-Customer-Driven-ebook/dp/B0979QYPWD/). There already exists a [package]('https://cran.r-project.org/web/packages/boot/index.html') dedicated to Bootstrap simulations, but it often requires more memory and computation than my personal laptop can manage and I find it somewhat cumbersome to use. I have built the `boot_ci()` function in the present package as a lightweight alternative with reasonable defaults for easy use within a Tidyverse workflow. In this vignette, I'll explain how to use it.  

### Passing a linear regression formula

The most common and general use case for `boot_ci()` is to build the confidence interval for a linear regression. Let's import the Boston housing price dataset from the MASS package:

```{r}
dat <- MASS::Boston
head(dat)

```

The goal is to estimate the median value of owner-occupied homes in an area ('medv' variable).

```{r}
formula <- "medv~nox+crim"
lm_fit <- lm(data = dat, formula = formula)
summary(lm_fit)

```
However, the standard error might be a poor estimate of the distribution of a coefficient if our data is not as nicely shaped and behaved as we'd like. Using `boot_ci()` allows us to Bootstrap confidence intervals for each coefficient.

```{r}
set.seed(1)
boot_ci(dat, formula)
```
Under the hood, `boot_ci()` relies heavily on assumptions and conventions for the sake of convenience and speed. For instance, it determines silently whether to use a linear or logistic regression based on the number of distinct values for the dependent variable. By default, it will run 100 Bootstrap simulations and returns the bounds of the 90% confidence interval, but you can override these values by explicitly passing the corresponding parameters.

```{r}
set.seed(1)
boot_ci(dat, formula, B = 200, conf.level = 0.80)
```

### Passing a logistic regression formula

If the dependent variable in your formula only has two values (TRUE/FALSE, Y/N, or 1/0), `boot_ci()` will automatically run a logistic regression. Let's look for example at data about biopsies of breast tumor, also from the MASS package

```{r}
biopsy_dat <- MASS::biopsy |>
  dplyr::select(-ID) |> # removing the ID column we won't use
  na.omit() |> # removing all rows with NA's
  mutate(class = factor(class, levels = c('benign', 'malignant'), labels = c(0, 1)))

# Standard model
log_mod <- glm("class~.", family= binomial(link = "logit"), data = biopsy_dat)
summary(log_mod)
```
Let's now build the corresponding Bootstrap confidence intervals.

```{r}
formula <- "class~."
boot_ci(biopsy_dat, formula, B = 400, conf.level = 0.95)
```




